---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RStanTVA

<!-- badges: start -->
[![R](https://github.com/mmrabe/RStanTVA/actions/workflows/rpkg.yml/badge.svg)](https://github.com/mmrabe/RStanTVA/actions/workflows/rpkg.yml)
<!-- badges: end -->

RStanTVA is an R package containing the StanTVA library and numerous convenience functions for generating, compiling, fitting, and analyzing (Stan)TVA models.

## Installation

You can install the development version of RStanTVA from [GitHub](https://github.com/mmrabe/RStanTVA) with:

``` r
remotes::install_github("mmrabe/RStanTVA")
```

## Example

Load the R package:

```{r example-load}
library(RStanTVA)
library(tidyverse)
```

Load example data from the parameter recovery study:

```{r example-data}
tva_data <- tva_recovery %>% filter(subject == 20)
tva_data
```


Generate a StanTVA model for partial report of 6 display locations with Gaussian $t_0$ and a free $K$ distribution:

```{r example-model, warning=FALSE}
library(RStanTVA)

tva_model <- stantva_model(
  location = 4,
  task = "pr",
  w_mode = "locations",
  t0_mode = "gaussian",
  K_mode = "free",
  sanity_checks = FALSE
)

tva_model

```


The generated Stan model looks like this:

``` stan
`r tva_model@code@code`
```


Fit `tva_model` to the `tva_data` using maximum-likelihood estimation (MLE):

```{r example-fit-mle}
tva_fit_mle <- optimizing(tva_model, tva_data)
tva_fit_mle$par[c("C","alpha","mu0","sigma0")]
```


Fit `tva_model` to the `tva_data` using maximum-likelihood estimation (MLE):

```{r example-fit}
tva_fit <- sampling(tva_model, tva_data)
tva_fit
```




