---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RStanTVA

<!-- badges: start -->
<!-- badges: end -->

RStanTVA is an R package containing the StanTVA library and numerous convenience functions for generating, compiling, fitting, and analyzing (Stan)TVA models.

## Installation

You can install the development version of RStanTVA from [GitHub](https://github.com/mmrabe/RStanTVA) with:

``` r
remotes::install_github("mmrabe/RStanTVA")
```

## Example

Load the R package:

```{r example-load}
library(RStanTVA)
library(tidyverse)
```

Load example data from the parameter recovery study:

```{r example-data}
tva_data <- tva_recovery %>% filter(subject == 20)
tva_data
```

Create a report and visualize the scores:

```{r example-report}
tva_data_report <- tva_report(tva_data)
tva_data_report


tva_data_report %>%
  group_by(exposure) %>%
  summarize(mean_score = mean(score)) %>%
  ggplot(aes(exposure, mean_score)) +
  theme_minimal() +
  geom_point(size = 0.2) +
  geom_line(linewidth = 0.2)
  


```

Generate a StanTVA model for partial report of 6 display locations with Gaussian $t_0$ and a free $K$ distribution:

```{r example-model}
library(RStanTVA)

tva_model <- stantva_model(
  location = 4,
  task = "pr",
  w_mode = "locations",
  t0_mode = "gaussian",
  K_mode = "free",
  sanity_checks = FALSE,
  priors = FALSE
)

tva_model

```


The generated Stan model looks like this:

``` stan
`r tva_model@code@code`
```


Fit `tva_model` to the `tva_data` using maximum-likelihood estimation (MLE):

```{r example-fit}
tva_fit <- optimizing(tva_model, tva_data)
str(tva_fit)
tva_fit$par[c("C","alpha","mu0","sigma0")]
```




